---
title: "Exp_1_analysis"
output: html_document
date: "2024-01-18"
---

### Data Analayis for Experiment 1 of Precommitment ###

# Basic R set up for analysis #
```{r setup, include=FALSE} 
### usefull packages ###
library("tidyr"); library("dplyr")
library("ggplot2"); library("data.table")
library("ggpubr") 
library("arm"); library("lme4")
library("reshape2"); library("Hmisc") 
library("psych"); library("corrplot")
## Set working directory ##
dataDir = '~/Documents/Github/precommitment/Exp_1_data'

```

# Load all the Data for analysis #
This will create the Clean Data list and the Final Data large data frame of all participants. 
If you want to subset to only look at the data frame of one particular subject the code looks like (CleanData$'101').
```{r load_data, include=FALSE} 
allFileNames <- list.files(path = dataDir)
allData <- lapply(allFileNames, function(fileName) {
  file <- fread(sprintf("%s/%s", dataDir, fileName)) # using fread for faster reading
  TFCols <- c("A_alwaysBest", "B_alwaysBest", "Best_changes", "Choice_Wait",
              "Choice_Commit", "Choice_Miss",  "Play_win_A",
              "Play_win_B", "Play_lost", "Play_lost_chosenA", "Play_lost_chosenB", "Precomm_B",
              "Precomm_A","Precomm_miss")
  TFNACols <- c("Press_hit", "Press_miss")
  file[, (TFCols) := lapply(.SD, function(x) as.numeric(as.logical(x))), .SDcols = TFCols]
  file[, (TFNACols) := lapply(.SD, function(x) as.numeric(as.logical(x))), .SDcols = TFNACols]
  return(file)
})
allDataCombined <- rbindlist(allData)

# Processing each file to include additional variables
CleanData <- lapply(allData, function(thisData) {
  trials <- 1:96
  thisData[, AverageValue := (OptionA + OptionB) / 2]
  thisData[, LargerOption := pmax(OptionA, OptionB)]
  thisData[, SmallerOption := pmin(OptionA, OptionB)]
  thisData[, centeredLargeOption := LargerOption - mean(LargerOption)]
  thisData[, centeredSmallOption := SmallerOption - mean(SmallerOption)]
  thisData[, centeredDifference := difference - 16]
  thisData[, centeredAverageValue := AverageValue - 50]
  thisData[, trial := trials - 48]
  return(thisData)
})
names(CleanData) <- unlist(lapply(allData, function(x) x$participant[1]))
CleanDataCombined <- rbindlist(CleanData)
FinalData <- CleanDataCombined[Choice_Miss == FALSE]

## Separate the data into the first and last blocks ##

# First block
firstData <- lapply(CleanData, function(thisFile) {
  firstPart <- thisFile[1:23, ]  # Extract the first 23 trials
  firstPart$block <- 0  # Add a column to indicate this is the first block
  return(firstPart)
})
firstDataCombined <- do.call("rbind", firstData)
firstClean <- firstDataCombined[firstDataCombined$Choice_Miss == FALSE, ]
# Last block
endData <- lapply(CleanData, function(thisFile) {
  endPart <- thisFile[73:95, ]  # Extract the last 23 trials
  endPart$block <- 1  # Add a column to indicate this is the last block
  return(endPart)
})
endDataCombined <- do.call("rbind", endData)
endClean <- endDataCombined[endDataCombined$Choice_Miss == FALSE, ]
# Combine first and last block
firstEndData <- do.call("rbind", c(firstData, endData))
firstEndClean <- firstEndData[firstEndData$Choice_Miss == FALSE, ]

```

# Check that success rate on the reaction time task waiting is approx. 50% #
```{r rt_task, echo=FALSE}
# Calculate overall success rate 
AverageHits <- mean(na.omit(FinalData$Press_hit), na.rm = TRUE)
sprintf('Overall success rate: %f', AverageHits)

# Create and order the data frame
df <- data.frame(Choice_Wait=FinalData$Choice_Wait, 
                 Press_hit=FinalData$Press_hit, 
                 participant=FinalData$participant)
df <- df[with(df, order(Choice_Wait, participant)), ]

# Calculate hit rates for each participant 
hitRates <- sapply(CleanData, function(thisData) {
  mean(na.omit(thisData$Press_hit), na.rm = TRUE)
})
hitRates <- hitRates[!is.na(hitRates)]

#visualize the hit rates
hitRatesdf <- data.frame(subj = seq_along(hitRates), hitRates)
hitRatesdf <- subset(hitRatesdf, hitRates != 0)
# Create the plot
ggplot(data = hitRatesdf, aes(x = subj, y = hitRates)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "Individuals", y = "Success Rates")

```

# Check on how many missed trials per participant #
```{r missed_data, echo=FALSE}
# Calculate the number of missed choices for each participant
MissingData <- lapply(CleanData, function(thisData) {
  data.frame(
    sub = unique(thisData$participant),
    numberMissed = sum(thisData$Choice_Miss, na.rm = TRUE)
  )
})

# Combine and order the data
MissingData <- do.call("rbind", MissingData)
MissingData <- MissingData[order(MissingData$numberMissed), ]

```

# Basic Desciptive Statistics of precommitment rates #
```{r descriptive, echo=FALSE}
# Average Precommit Total
pCommitTotal <- mean(FinalData$Choice_Commit, na.rm = TRUE)
sdCommitTotal <- sd(FinalData$Choice_Commit, na.rm = TRUE)

# Average Initial Value 
avValue <- mean(FinalData$AverageValue, na.rm = TRUE)
sdValue <- sd(FinalData$AverageValue, na.rm = TRUE)

# Table of by difference 
differenceSummary <- FinalData %>%
  group_by(difference) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))
print(differenceSummary)

pCommitSub <- FinalData %>%
  group_by(participant) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

sortpCommit <- round(sort(pCommitSub$pCommit), digits = 2)

# Wilcox test to see if different from optimal behavior 
wilcoxPcommit <- wilcox.test(pCommitSub$pCommit, mu = 0.5, alternative = "two.sided")
print(wilcoxPcommit)
```

# Differences between the initial value differences #
```{r differences, echo=FALSE}

four = FinalData[FinalData$difference == "4",]
twelve  = FinalData[FinalData$difference == "12",]
twenty = FinalData[FinalData$difference == "20",]
twentyEight= FinalData[FinalData$difference == "28",]
Small = FinalData[FinalData$difference == "4" | FinalData$difference == "12",]
Large =  FinalData[FinalData$difference == "20" | FinalData$difference == "28",]
#Create the data frames
pCommitFour = data.frame(summarise(group_by(four, participant) , pCommit = mean(Choice_Commit)))
pCommitTwelve = data.frame(summarise(group_by(twelve, participant) , pCommit = mean(Choice_Commit)))
pCommitTwenty = data.frame(summarise(group_by(twenty, participant) , pCommit = mean(Choice_Commit)))
pCommitTwentyEight = data.frame(summarise(group_by(twentyEight, participant) , pCommit = mean(Choice_Commit)))
pCommitSmall = data.frame(summarise(group_by(Small, participant) , pCommit = mean(Choice_Commit)))
pCommitLarge = data.frame(summarise(group_by(Large, participant) , pCommit = mean(Choice_Commit)))

#Wilcox test of differences 
wilcoxSmaller = wilcox.test(pCommitTwelve$pCommit, pCommitFour$pCommit, paired = TRUE)
wilcoxLarger = wilcox.test(pCommitTwentyEight$pCommit, pCommitTwenty$pCommit, paired = TRUE)
wilcoxMiddle = wilcox.test(pCommitTwenty$pCommit, pCommitTwelve$pCommit, paired = TRUE)
wilcoxLarge.Small = wilcox.test(pCommitLarge$pCommit, pCommitSmall$pCommit, paired= TRUE)

# box-plot graph of pCommit by difference 
df_box = data.frame(summarise(group_by(FinalData, participant, difference), pCommit=mean(Choice_Commit)))
ggplot(data=df_box, aes(x=difference, y=pCommit, group=difference)) +
  geom_boxplot(fill='#A4A4A4', color="black", outlier.colour="red")+
  theme_minimal() + 
  labs(x="Initial Differences", y="Proportion Precommit", title="")

```

# Deviation from optimal #
```{r optimal, echo=FALSE}

# Summarization
pCommitSumm <- FinalData %>%
  group_by(participant) %>%
  summarise(sumCommit = sum(Choice_Commit, na.rm = TRUE))

sumCommit <- sum(pCommitSumm$sumCommit, na.rm = TRUE)

# Chi-square test
nTrialsTotal <- 6057
diff <- sumCommit / nTrialsTotal
optimalitytest <- binom.test(sumCommit, nTrialsTotal, p = 0.5, alternative = "two.sided")

# Confidence Interval
p <- 0.5
UB <- diff + (1.64 * sqrt((p * (1 - p)) / nTrialsTotal))
LB <- diff - (1.64 * sqrt((p * (1 - p)) / nTrialsTotal))

# Create the optimal df
diff_opt <- c(4, 12, 20, 28)
opt <- c(0, 0, 1, 1)
df_opt <- data.frame(diff = diff_opt, opt = opt)

# Group mean and median pCommit by differences
df_mean <- FinalData %>%
  group_by(difference) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

df_graph <- FinalData %>%
  group_by(participant, difference) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

df_median <- df_graph %>%
  group_by(difference) %>%
  summarise(pCommit = median(pCommit, na.rm = TRUE))

# Plotting
ggplot() +
  geom_line(data = df_graph, aes(x = difference, y = pCommit, group = participant), color = "gray") +
  geom_point(data = df_graph, aes(x = difference, y = pCommit, group = participant), color = "gray") +
  geom_line(data = df_opt, aes(x = diff, y = opt), color = "red", size = 1) +
  geom_point(data = df_opt, aes(x = diff, y = opt), color = "red", size = 3) +
  geom_line(data = df_median, aes(x = difference, y = pCommit), size = 1) +
  geom_point(data = df_median, aes(x = difference, y = pCommit), size = 3) +
  scale_color_brewer(palette = "Paired") +
  theme_minimal() +
  labs(x = "Initial Differences", y = "Proportion Precommit", title = "Proportion Precommit by Initial Differences")

```

# Average Value #
```{r average_value, echo=FALSE}
# Average value = average of two initial values
df_val <- FinalData %>%
  group_by(AverageValue) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

averageGLM <- glm(pCommit ~ AverageValue, data = df_val)

# Plot with geom_smooth (which fits a smoother to the data)
ggplot(df_val, aes(x = AverageValue, y = pCommit)) +
  geom_smooth(color = "black") +
  theme_classic() +
  labs(x = "Average Value", y = "Proportion Precommit", title = "Proportion of Precommit by Average Value")

# Plot with geom_point
ggplot(df_val, aes(x = AverageValue, y = pCommit)) +
  geom_point() +
  theme_classic() +
  labs(x = "Average Value", y = "Proportion Precommit", title = "Proportion of Precommit by Average Value")

```

# Learning # 
```{r learning, echo=FALSE}

learning <- FinalData %>%
  group_by(trial) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

# Plotting
ggplot(learning, aes(x = trial, y = pCommit)) +
  geom_smooth(color = "black") +
  theme_minimal() +
  labs(x = "Trial Number", y = "Proportion Precommit", title = "")

ggplot(learning, aes(x = trial, y = pCommit)) +
  geom_point() +
  theme_classic() +
  labs(x = "Trial Number", y = "Proportion Precommit", title = "Learning")

# Data summarization for T-tests
summarize_by_difference <- function(data, difference) {
  data %>%
    filter(difference == difference) %>%
    group_by(participant) %>%
    summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))
}

# Applying the summarization function
pCommitFirst4 <- summarize_by_difference(firstClean, 4)
pCommitFirst12 <- summarize_by_difference(firstClean, 12)
pCommitFirst20 <- summarize_by_difference(firstClean, 20)
pCommitFirst28 <- summarize_by_difference(firstClean, 28)

pCommitEnd4 <- summarize_by_difference(endClean, 4)
pCommitEnd12 <- summarize_by_difference(endClean, 12)
pCommitEnd20 <- summarize_by_difference(endClean, 20)
pCommitEnd28 <- summarize_by_difference(endClean, 28)

# T-tests
testFour <- t.test(pCommitEnd4$pCommit, pCommitFirst4$pCommit, paired = TRUE)
testTwelve <- t.test(pCommitEnd12$pCommit, pCommitFirst12$pCommit, paired = TRUE)
testTwenty <- t.test(pCommitEnd20$pCommit, pCommitFirst20$pCommit, paired = TRUE)
testTwentyEight <- t.test(pCommitEnd28$pCommit, pCommitFirst28$pCommit, paired = TRUE)

# GLMMs
firstDiffModel <- glmer(Choice_Commit ~ centeredDifference + (1 | participant), data = firstClean, family = binomial, nAGQ = 0)
endDiffModel <- glmer(Choice_Commit ~ centeredDifference + (1 | participant), data = endClean, family = binomial, nAGQ = 0)

firstValueModel <- glmer(Choice_Commit ~ centeredAverageValue + (1 | participant), data = firstClean, family = binomial, nAGQ = 0)
endValueModel <- glmer(Choice_Commit ~ centeredAverageValue + (1 | participant), data = endClean, family = binomial, nAGQ = 0)

## Learning Graphs ##
#Seperated into four different plots 
four = FinalData[FinalData$difference == "4",]
twelve  = FinalData[FinalData$difference == "12",]
twenty = FinalData[FinalData$difference == "20",]
twentyEight= FinalData[FinalData$difference == "28",]

learnFour = data.frame(summarise(group_by(four, trial), pCommit=mean(Choice_Commit)))
learnTwelve = data.frame(summarise(group_by(twelve, trial), pCommit=mean(Choice_Commit)))
learnTwenty = data.frame(summarise(group_by(twenty, trial), pCommit=mean(Choice_Commit)))
learnTwentyEight = data.frame(summarise(group_by(twentyEight, trial), pCommit=mean(Choice_Commit)))

#Point plots 
#Four
ggplot()+
  geom_point(data=learnFour, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Four")
#Twelve
ggplot()+
  geom_point(data=learnTwelve, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twelve")
#Twenty
ggplot()+
  geom_point(data=learnTwenty, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twenty")
#TwentyEight
ggplot()+
  geom_point(data=learnTwentyEight, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="TwentyEight")
#Smooth trend plots 
#Four
ggplot()+
  geom_smooth(data=learnFour, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Four")
#Twelve
ggplot()+
  geom_smooth(data=learnTwelve, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twelve")
#Twenty
ggplot()+
  geom_smooth(data=learnTwenty, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twenty")
#TwentyEight
ggplot()+
  geom_smooth(data=learnTwentyEight, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="TwentyEight")


```
