---
title: "Exp_1_analysis"
output: html_document
date: "2024-01-18"
---

### Data Analayis for Experiment 1 of Precommitment ###

# Basic R set up for analysis #
```{r setup, include=FALSE} 
### usefull packages ###
library("tidyr"); library("dplyr")
library("ggplot2"); library("data.table")
library("ggpubr") 
library("arm"); library("lme4")
library("reshape2"); library("Hmisc") 
library("psych"); library("corrplot")
## Set working directory ##
dataDir = '~/Documents/Github/precommitment/Exp_1_data'

```

# Load all the Data for analysis #
This will create the Clean Data list and the Final Data large data frame of all participants. 
If you want to subset to only look at the data frame of one particular subject the code looks like (CleanData$'101').
```{r load_data, include=FALSE} 
allFileNames <- list.files(path = dataDir)
allData <- lapply(allFileNames, function(fileName) {
  file <- fread(sprintf("%s/%s", dataDir, fileName)) # using fread for faster reading
  TFCols <- c("A_alwaysBest", "B_alwaysBest", "Best_changes", "Choice_Wait",
              "Choice_Commit", "Choice_Miss",  "Play_win_A",
              "Play_win_B", "Play_lost", "Play_lost_chosenA", "Play_lost_chosenB", "Precomm_B",
              "Precomm_A","Precomm_miss")
  TFNACols <- c("Press_hit", "Press_miss")
  file[, (TFCols) := lapply(.SD, function(x) as.numeric(as.logical(x))), .SDcols = TFCols]
  file[, (TFNACols) := lapply(.SD, function(x) as.numeric(as.logical(x))), .SDcols = TFNACols]
  return(file)
})
allDataCombined <- rbindlist(allData)

# Processing each file to include additional variables
CleanData <- lapply(allData, function(thisData) {
  trials <- 1:96
  thisData[, AverageValue := (OptionA + OptionB) / 2]
  thisData[, LargerOption := pmax(OptionA, OptionB)]
  thisData[, SmallerOption := pmin(OptionA, OptionB)]
  thisData[, centeredLargeOption := LargerOption - mean(LargerOption)]
  thisData[, centeredSmallOption := SmallerOption - mean(SmallerOption)]
  thisData[, centeredDifference := difference - 16]
  thisData[, centeredAverageValue := AverageValue - 50]
  thisData[, trial := trials - 48]
  return(thisData)
})
names(CleanData) <- unlist(lapply(allData, function(x) x$participant[1]))
CleanDataCombined <- rbindlist(CleanData)
FinalData <- CleanDataCombined[Choice_Miss == FALSE]

## Separate the data into the first and last blocks ##

# First block
firstData <- lapply(CleanData, function(thisFile) {
  firstPart <- thisFile[1:23, ]  # Extract the first 23 trials
  firstPart$block <- 0  # Add a column to indicate this is the first block
  return(firstPart)
})
firstDataCombined <- do.call("rbind", firstData)
firstClean <- firstDataCombined[firstDataCombined$Choice_Miss == FALSE, ]
# Last block
endData <- lapply(CleanData, function(thisFile) {
  endPart <- thisFile[73:95, ]  # Extract the last 23 trials
  endPart$block <- 1  # Add a column to indicate this is the last block
  return(endPart)
})
endDataCombined <- do.call("rbind", endData)
endClean <- endDataCombined[endDataCombined$Choice_Miss == FALSE, ]
# Combine first and last block
firstEndData <- do.call("rbind", c(firstData, endData))
firstEndClean <- firstEndData[firstEndData$Choice_Miss == FALSE, ]

```

# Check that success rate on the reaction time task waiting is approx. 50% #
```{r rt_task, echo=FALSE}
# Calculate overall success rate 
AverageHits <- mean(na.omit(FinalData$Press_hit), na.rm = TRUE)
sprintf('Overall success rate: %f', AverageHits)

# Create and order the data frame
df <- data.frame(Choice_Wait=FinalData$Choice_Wait, 
                 Press_hit=FinalData$Press_hit, 
                 participant=FinalData$participant)
df <- df[with(df, order(Choice_Wait, participant)), ]

# Calculate hit rates for each participant 
hitRates <- sapply(CleanData, function(thisData) {
  mean(na.omit(thisData$Press_hit), na.rm = TRUE)
})
hitRates <- hitRates[!is.na(hitRates)]

#visualize the hit rates
hitRatesdf <- data.frame(subj = seq_along(hitRates), hitRates)
hitRatesdf <- subset(hitRatesdf, hitRates != 0)
# Create the plot
ggplot(data = hitRatesdf, aes(x = subj, y = hitRates)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x = "Individuals", y = "Success Rates")

```

# Check on how many missed trials per participant #
```{r missed_data, echo=FALSE}
# Calculate the number of missed choices for each participant
MissingData <- lapply(CleanData, function(thisData) {
  data.frame(
    sub = unique(thisData$participant),
    numberMissed = sum(thisData$Choice_Miss, na.rm = TRUE)
  )
})

# Combine and order the data
MissingData <- do.call("rbind", MissingData)
MissingData <- MissingData[order(MissingData$numberMissed), ]

```

# Basic Desciptive Statistics of precommitment rates #
```{r descriptive, echo=FALSE}
# Average Precommit Total
pCommitTotal <- mean(FinalData$Choice_Commit, na.rm = TRUE)
sdCommitTotal <- sd(FinalData$Choice_Commit, na.rm = TRUE)

# Average Initial Value 
avValue <- mean(FinalData$AverageValue, na.rm = TRUE)
sdValue <- sd(FinalData$AverageValue, na.rm = TRUE)

# Table of by difference 
differenceSummary <- FinalData %>%
  group_by(difference) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))
print(differenceSummary)

pCommitSub <- FinalData %>%
  group_by(participant) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

sortpCommit <- round(sort(pCommitSub$pCommit), digits = 2)

# Wilcox test to see if different from optimal behavior 
wilcoxPcommit <- wilcox.test(pCommitSub$pCommit, mu = 0.5, alternative = "two.sided")
print(wilcoxPcommit)
```

# Differences between the initial value differences #
```{r differences, echo=FALSE}

four = FinalData[FinalData$difference == "4",]
twelve  = FinalData[FinalData$difference == "12",]
twenty = FinalData[FinalData$difference == "20",]
twentyEight= FinalData[FinalData$difference == "28",]
Small = FinalData[FinalData$difference == "4" | FinalData$difference == "12",]
Large =  FinalData[FinalData$difference == "20" | FinalData$difference == "28",]
#Create the data frames
pCommitFour = data.frame(summarise(group_by(four, participant) , pCommit = mean(Choice_Commit)))
pCommitTwelve = data.frame(summarise(group_by(twelve, participant) , pCommit = mean(Choice_Commit)))
pCommitTwenty = data.frame(summarise(group_by(twenty, participant) , pCommit = mean(Choice_Commit)))
pCommitTwentyEight = data.frame(summarise(group_by(twentyEight, participant) , pCommit = mean(Choice_Commit)))
pCommitSmall = data.frame(summarise(group_by(Small, participant) , pCommit = mean(Choice_Commit)))
pCommitLarge = data.frame(summarise(group_by(Large, participant) , pCommit = mean(Choice_Commit)))

#Wilcox test of differences 
wilcoxSmaller = wilcox.test(pCommitTwelve$pCommit, pCommitFour$pCommit, paired = TRUE)
wilcoxLarger = wilcox.test(pCommitTwentyEight$pCommit, pCommitTwenty$pCommit, paired = TRUE)
wilcoxMiddle = wilcox.test(pCommitTwenty$pCommit, pCommitTwelve$pCommit, paired = TRUE)
wilcoxLarge.Small = wilcox.test(pCommitLarge$pCommit, pCommitSmall$pCommit, paired= TRUE)

# box-plot graph of pCommit by difference 
df_box = data.frame(summarise(group_by(FinalData, participant, difference), pCommit=mean(Choice_Commit)))
ggplot(data=df_box, aes(x=difference, y=pCommit, group=difference)) +
  geom_boxplot(fill='#A4A4A4', color="black", outlier.colour="red")+
  theme_minimal() + 
  labs(x="Initial Differences", y="Proportion Precommit", title="")

```

# Deviation from optimal #
```{r optimal, echo=FALSE}

# Summarization
pCommitSumm <- FinalData %>%
  group_by(participant) %>%
  summarise(sumCommit = sum(Choice_Commit, na.rm = TRUE))

sumCommit <- sum(pCommitSumm$sumCommit, na.rm = TRUE)

# Chi-square test
nTrialsTotal <- 6057
diff <- sumCommit / nTrialsTotal
optimalitytest <- binom.test(sumCommit, nTrialsTotal, p = 0.5, alternative = "two.sided")

# Confidence Interval
p <- 0.5
UB <- diff + (1.64 * sqrt((p * (1 - p)) / nTrialsTotal))
LB <- diff - (1.64 * sqrt((p * (1 - p)) / nTrialsTotal))

# Create the optimal df
diff_opt <- c(4, 12, 20, 28)
opt <- c(0, 0, 1, 1)
df_opt <- data.frame(diff = diff_opt, opt = opt)

# Group mean and median pCommit by differences
df_mean <- FinalData %>%
  group_by(difference) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

df_graph <- FinalData %>%
  group_by(participant, difference) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

df_median <- df_graph %>%
  group_by(difference) %>%
  summarise(pCommit = median(pCommit, na.rm = TRUE))

# Plotting
ggplot() +
  geom_line(data = df_graph, aes(x = difference, y = pCommit, group = participant), color = "gray") +
  geom_point(data = df_graph, aes(x = difference, y = pCommit, group = participant), color = "gray") +
  geom_line(data = df_opt, aes(x = diff, y = opt), color = "red", size = 1) +
  geom_point(data = df_opt, aes(x = diff, y = opt), color = "red", size = 3) +
  geom_line(data = df_median, aes(x = difference, y = pCommit), size = 1) +
  geom_point(data = df_median, aes(x = difference, y = pCommit), size = 3) +
  scale_color_brewer(palette = "Paired") +
  theme_minimal() +
  labs(x = "Initial Differences", y = "Proportion Precommit", title = "Proportion Precommit by Initial Differences")

```

# Average Value #
```{r average_value, echo=FALSE}
# Average value = average of two initial values
df_val <- FinalData %>%
  group_by(AverageValue) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

averageGLM <- glm(pCommit ~ AverageValue, data = df_val)

# Plot with geom_smooth (which fits a smoother to the data)
ggplot(df_val, aes(x = AverageValue, y = pCommit)) +
  geom_smooth(color = "black") +
  theme_classic() +
  labs(x = "Average Value", y = "Proportion Precommit", title = "Proportion of Precommit by Average Value")

# Plot with geom_point
ggplot(df_val, aes(x = AverageValue, y = pCommit)) +
  geom_point() +
  theme_classic() +
  labs(x = "Average Value", y = "Proportion Precommit", title = "Proportion of Precommit by Average Value")

```

# Learning # 
```{r learning, echo=FALSE}

learning <- FinalData %>%
  group_by(trial) %>%
  summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

# Plotting
ggplot(learning, aes(x = trial, y = pCommit)) +
  geom_smooth(color = "black") +
  theme_minimal() +
  labs(x = "Trial Number", y = "Proportion Precommit", title = "")

ggplot(learning, aes(x = trial, y = pCommit)) +
  geom_point() +
  theme_classic() +
  labs(x = "Trial Number", y = "Proportion Precommit", title = "Learning")

# Data summarization for T-tests
summarize_by_difference <- function(data, difference) {
  data %>%
    filter(difference == difference) %>%
    group_by(participant) %>%
    summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))
}

# Applying the summarization function
pCommitFirst4 <- summarize_by_difference(firstClean, 4)
pCommitFirst12 <- summarize_by_difference(firstClean, 12)
pCommitFirst20 <- summarize_by_difference(firstClean, 20)
pCommitFirst28 <- summarize_by_difference(firstClean, 28)

pCommitEnd4 <- summarize_by_difference(endClean, 4)
pCommitEnd12 <- summarize_by_difference(endClean, 12)
pCommitEnd20 <- summarize_by_difference(endClean, 20)
pCommitEnd28 <- summarize_by_difference(endClean, 28)

# T-tests
testFour <- t.test(pCommitEnd4$pCommit, pCommitFirst4$pCommit, paired = TRUE)
testTwelve <- t.test(pCommitEnd12$pCommit, pCommitFirst12$pCommit, paired = TRUE)
testTwenty <- t.test(pCommitEnd20$pCommit, pCommitFirst20$pCommit, paired = TRUE)
testTwentyEight <- t.test(pCommitEnd28$pCommit, pCommitFirst28$pCommit, paired = TRUE)

# GLMMs
firstDiffModel <- glmer(Choice_Commit ~ centeredDifference + (1 | participant), data = firstClean, family = binomial, nAGQ = 0)
endDiffModel <- glmer(Choice_Commit ~ centeredDifference + (1 | participant), data = endClean, family = binomial, nAGQ = 0)

firstValueModel <- glmer(Choice_Commit ~ centeredAverageValue + (1 | participant), data = firstClean, family = binomial, nAGQ = 0)
endValueModel <- glmer(Choice_Commit ~ centeredAverageValue + (1 | participant), data = endClean, family = binomial, nAGQ = 0)

## Learning Graphs ##
#Seperated into four different plots 
four = FinalData[FinalData$difference == "4",]
twelve  = FinalData[FinalData$difference == "12",]
twenty = FinalData[FinalData$difference == "20",]
twentyEight= FinalData[FinalData$difference == "28",]

learnFour = data.frame(summarise(group_by(four, trial), pCommit=mean(Choice_Commit)))
learnTwelve = data.frame(summarise(group_by(twelve, trial), pCommit=mean(Choice_Commit)))
learnTwenty = data.frame(summarise(group_by(twenty, trial), pCommit=mean(Choice_Commit)))
learnTwentyEight = data.frame(summarise(group_by(twentyEight, trial), pCommit=mean(Choice_Commit)))

#Point plots 
#Four
ggplot()+
  geom_point(data=learnFour, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Four")
#Twelve
ggplot()+
  geom_point(data=learnTwelve, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twelve")
#Twenty
ggplot()+
  geom_point(data=learnTwenty, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twenty")
#TwentyEight
ggplot()+
  geom_point(data=learnTwentyEight, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="TwentyEight")
#Smooth trend plots 
#Four
ggplot()+
  geom_smooth(data=learnFour, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Four")
#Twelve
ggplot()+
  geom_smooth(data=learnTwelve, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twelve")
#Twenty
ggplot()+
  geom_smooth(data=learnTwenty, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="Twenty")
#TwentyEight
ggplot()+
  geom_smooth(data=learnTwentyEight, aes(x=trial, y=pCommit))+
  theme_minimal() + labs(x="Trial Number", y="Proportion Precommit", title="TwentyEight")


```


# Models # 
```{r models, echo=FALSE}

# Model of initial value difference 
# This model assesses the effect of Difference on Choice_Commit
diffModel <- glmer(Choice_Commit ~ centeredDifference + (1 | participant), 
                   data = FinalData, family = binomial, nAGQ = 0)

# Model of average initial value
# This model assesses the effect of Average Value on Choice_Commit
valueModel <- glmer(Choice_Commit ~ centeredAverageValue + (1 | participant), 
                    data = FinalData, family = binomial, nAGQ = 0)

# Model of trial number 
# This model assesses the effect of trial on Choice_Commit
trialModel <- glmer(Choice_Commit ~ trial + (1 | participant), 
                    data = FinalData, family = binomial, nAGQ = 0)

# Interaction Model 
# This model assesses the interaction between Average Value and Difference
interactionModel <- glmer(Choice_Commit ~ centeredAverageValue + centeredDifference + 
                          centeredAverageValue * centeredDifference + (1 | participant), 
                          data = FinalData, family = binomial, nAGQ = 0)

# Full Model 
# This model includes all main effects and interactions with trial
fullModel <- glmer(Choice_Commit ~ centeredAverageValue + centeredDifference + trial + 
                   centeredAverageValue * trial + centeredDifference * trial + (1 | participant), 
                   data = FinalData, family = binomial, nAGQ = 0)

# Model Comparison 
# Comparing all models
modelTest <- anova(diffModel, valueModel, trialModel, interactionModel, fullModel) 

# Comparing Interaction Model with Full Model
betterTest <- anova(interactionModel, fullModel) 


```

# Individual participant coefficients #
```{r ind_coef, echo = FALSE}
# Get the individual beta coefficients for correlations with questionniare measures
indBetas = list()
for (i in 1: length(CleanData)){
  d = CleanData[[i]]
  par = unique(d$participant)
  model = lm(Choice_Commit ~ centeredAverageValue * trial * centeredDifference, data=d)
  coef = data.frame(model$coefficients)
  coef = data.frame(t(coef), par)
  indBetas[[i]] = coef
}
allindBetas = do.call("rbind", indBetas)

```

# Load questionnaire data #
```{r load_questionnaires, echo = FALSE}
## Set directory to get questionnaire data ##
questionnaire = read.csv('~/Documents/Github/precommitment/Exp_1_questionnaire.csv')

## Participant ID variable ##
id <- questionnaire$Q1
```

# BIS-11 variables #
``` {r BIS-11, echo=FALSE}

# Load data for BIS-11
bisdata <- questionnaire[, 31:60]

# Fix reverse coding on the data
keys <- c(-1, 1, 1, 1, 1, 1, -1, -1, -1, -1, 1, -1, -1, 1, -1, 1, 1, 1, 1, -1, 1, 1, 1, 1, 1, 1, 1, 1, -1, -1)
clean_bis <- reverse.code(keys, bisdata, mini = 1, maxi = 4)

# Calculate the full sum of the BIS
fullsum <- rowSums(clean_bis)
participantBisSum <- data.frame(id, fullsum)

# 2nd order factors of BIS
# Attentional factor
attentional_indices <- c(5, 9, 11, 20, 28, 6, 24, 26)
attentional <- clean_bis[, attentional_indices]
attentionalSum <- rowSums(attentional)
participantAttentional <- data.frame(id, attentionalSum)
Attentional <- participantAttentional[order(participantAttentional$id), "attentionalSum"]

# Motor factor
motor_indices <- c(2, 3, 4, 17, 19, 22, 25, 16, 21, 23, 30)
motor <- clean_bis[, motor_indices]
motorSum <- rowSums(motor)
participantMotor <- data.frame(id, motorSum)
Motor <- participantMotor[order(participantMotor$id), "motorSum"]

# Nonplanning factor
nonplanning_indices <- c(1, 7, 8, 12, 13, 14, 10, 15, 18, 27, 29)
nonplanning <- clean_bis[, nonplanning_indices]
nonplanningSum <- rowSums(nonplanning)
participantNonplanning <- data.frame(id, nonplanningSum)
Nonplanning <- participantNonplanning[order(participantNonplanning$id), "nonplanningSum"]

# Sort the main BIS data.frame
sortBIS <- participantBisSum[order(participantBisSum$id),]

# Combine full sum and 2nd order factors into one data frame
participantBis <- data.frame(
  FullSum = sortBIS$fullsum,
  Attentional,
  Motor,
  Nonplanning
)

```

# K-score calculation from MCQ #
``` {r K-score, echo=FALSE}
#K score calculation 
rawdd <- questionnaire[,3:29]
#remove repeated question
ddminus <- rawdd[,-c(26)]
#today = 0, later = 1
ddfixed <- lapply(ddminus, function(x) x-1)
#d <- data.frame(ddfixed)
ddDF <- data.frame(id, ddfixed)
#sort the data.frames
sortdd <- ddDF[order(ddDF$id),]
#remove the participant number from the dataframe 
noIDdd <- data.frame(sortdd[, -c(1)])
#Variables from Questionaire 
Vi <- c(54, 55, 19, 31, 14, 47, 15, 25, 78, 40, 11, 67, 34, 27, 69, 49, 80, 24 ,33, 28, 34, 25, 41, 54, 54, 22)
Vd <- c(55, 75, 25, 85, 25, 50, 35, 60, 80, 55, 30, 75, 35, 50, 85, 60, 85, 35, 80, 30, 50, 30, 75, 60, 80, 25)
values <- data.frame(Vi, Vd)
days <- c(117, 61, 53, 7, 19, 160, 13, 14, 162, 62, 7, 119, 186, 21, 91, 89, 157, 29, 14, 179, 30, 80, 20, 111, 30, 136)
t <- as.numeric(days)
valueDiff <- lapply(values, function(x) Vi/Vd)
R <- valueDiff$Vi
Rcoeff <- (1 - (1/R))
dfKscore <- rbind(Rcoeff, t, noIDdd)
ddData <- data.frame(t(dfKscore))
#X1 = Rcoeff, X2 = time
kscore <- vector(length=65)
#loop to acquire each participants k score
for(i in 1: 65){
  choices <- noIDdd[i,]
  bind <- rbind(Rcoeff=Rcoeff, t=t, choices=choices)
  df <- data.frame(t(bind))
  LR <- glm(choices ~ -1 + Rcoeff + t, family ="binomial", data=df)
  k = LR$coefficients[2] / LR$coefficients[1]
  kscore[i] = k
  #Fix participant who always waited to 0.00016
  kscore <- replace(kscore, kscore<0, 0.00016)
}
logkscore = log(kscore)

```

# Questionnaire correlations #
``` {r correlations, echo=FALSE}
# Sort pCommitSub by participant
pCommitSub <- pCommitSub[order(pCommitSub$participant),]

# Create a data frame with necessary variables
questionnaire_df <- data.frame(
  logkscore,
  FullSum = participantBis$FullSum,
  pCommit = pCommitSub$pCommit,
  allindBetas
)

# Calculate correlation matrix
questionnaire_COR <- cor(questionnaire_df, use = "complete.obs")

# Calcualte the significnce of the correlations 
# Initialize empty matrices p-values
p_matrix <- matrix(NA, ncol = ncol(questionnaire_df), nrow = ncol(questionnaire_df),
                   dimnames = list(names(questionnaire_df), names(questionnaire_df)))

# Compute correlations and p-values
for (i in 1:ncol(questionnaire_df)) {
  for (j in 1:ncol(questionnaire_df)) {
    temp <- cor.test(questionnaire_df[, i], questionnaire_df[, j], method = "pearson")
    p_matrix[i, j] <- temp$p.value
  }
}

# View the p-value matrix
p_matrix

# Define color palette for correlation plots
col2 <- colorRampPalette(c(
  "#053061", "#2166AC", "#4393C3", "#92C5DE", 
  "#D1E5F0", "#FFFFFF", "#FDDBC7", "#F4A582", 
  "#D6604D", "#B2182B", "#67001F"
))

# Correlation plot 
corrplot(
  questionnaire_COR, 
  is.corr = FALSE, 
  type = "upper", 
  outline = TRUE, 
  insig = "p-value", 
  sig.level = 0.05, # Adjusted significance level
  na.label = "NA", 
  na.label.col = "grey", 
  title = "Correlation Matrix", 
  method = "color", 
  tl.col = "black", 
  tl.cex = 0.8, 
  col = col2(200)
)

```


