---
title: "Exp_2_analysis"
output: html_document
date: "2024-01-19"
---

### Data Analayis for Experiment 2 of Precommitment ###

# Basic R set up for analysis #
```{r setup, echo=FALSE} 
### usefull packages ###
library("tidyr"); library("dplyr")
library("ggplot2"); library("data.table")
library("ggpubr"); library("purrr")
library("arm"); library("lme4")
library("reshape2"); library("Hmisc") 
library("psych"); library("corrplot")
## Set working directory ##
dataDir = '~/Documents/Github/precommitment/Exp_2_data'
```

# Load all the Data for analysis #

This portion will create the Clean Data list and the Final Data large dataframe of all participants. 
If you want to subset to only look at the dataframe of one particular subject the code looks like (CleanData\$'1001').
```{r load data, echo=FALSE}
allFileNames <- list.files(path = dataDir)
allData <- lapply(allFileNames, function(fileName) {
  file <- fread(sprintf("%s/%s", dataDir, fileName)) # using fread for faster reading
  TFCols <- c("Choice_Wait", "Choice_Commit", "Choice_Miss", "Play_win_A", 
              "Play_win_B", "Play_lost", "Play_lost_chosenA", "Play_lost_chosenB", 
              "Precomm_B", "Precomm_A","Precomm_miss")
  TFNACols <- c("Press_hit", "Press_miss")
  file[, (TFCols) := lapply(.SD, function(x) as.numeric(as.logical(x))), .SDcols = TFCols]
  file[, (TFNACols) := lapply(.SD, function(x) as.numeric(as.logical(x))), .SDcols = TFNACols]
  return(file)
})
allDataCombined <- rbindlist(allData)

# Processing each file to include additional variables
CleanData <- lapply(allData, function(thisData) {
  trials <- 1:96
  thisData[, AverageValue := (OptionA + OptionB) / 2]
  thisData[, LargerOption := pmax(OptionA, OptionB)]
  thisData[, SmallerOption := pmin(OptionA, OptionB)]
  thisData[, centeredLargeOption := LargerOption - mean(LargerOption)]
  thisData[, centeredSmallOption := SmallerOption - mean(SmallerOption)]
  thisData[, centeredDifference := difference - 16]
  thisData[, centeredAverageValue := AverageValue - 50]
  thisData[, trial := trials - 48]
  return(thisData)
})
names(CleanData) <- unlist(lapply(allData, function(x) x$participant[1]))
CleanDataCombined <- rbindlist(CleanData)
FinalData <- CleanDataCombined[Choice_Miss == FALSE]

```

# Seperate the Data into a Stable and Volatile List #
This portion will create two separate lists (volatileList and stableList) and data frames (volatileData and stableData)
```{r group data, echo=FALSE}

# Function to filter by condition and ensure Choice_Miss is logical
filter_by_condition <- function(df, condition) {
  df <- df %>% filter(Condition[1] == condition)

  # Ensure Choice_Miss is a logical column
  if(!is.logical(df$Choice_Miss)) {
    df$Choice_Miss <- as.logical(df$Choice_Miss)
  }
  return(df)
}

# Apply the function to each dataframe in the list but keep as a list
volatileList <- lapply(CleanData, filter_by_condition, condition = "Volatile")
stableList <- lapply(CleanData, filter_by_condition, condition = "Stable")

# Remove empty data frames from the lists
volatileList <- volatileList[sapply(volatileList, nrow) > 0]
stableList <- stableList[sapply(stableList, nrow) > 0]


# Apply the function to each dataframe in the list and bind rows
volatileGroup <- do.call(rbind, lapply(CleanData, filter_by_condition, condition = "Volatile"))
stableGroup <- do.call(rbind, lapply(CleanData, filter_by_condition, condition = "Stable"))

# Filter out Choice_Miss == TRUE
volatileData <- volatileGroup %>% filter(!Choice_Miss)
stableData <- stableGroup %>% filter(!Choice_Miss)

```

# Check on how many missed trials per participant #
The largest missing data participants over 10%: Stable: 1083(12); 1028(18); 1182(20) Volatile: 1166(15); 1147(17)
```{r missing, echo=FALSE}

# Function to calculate missing data for each group
calculate_missing_data <- function(data_group) {
  map_dfr(data_group, function(df) {
    data.frame(
      participant = unique(df$participant),
      numberMissed = sum(as.logical(df$Choice_Miss), na.rm = TRUE)
    )
  }) %>%
  arrange(numberMissed) %>%
  mutate(averageMissing = mean(numberMissed))
}

# Calculate missing data for all participants, stable, and volatile conditions
allMissingData <- calculate_missing_data(CleanData)
stableMissingData <- calculate_missing_data(stableList)
volatileMissingData <- calculate_missing_data(volatileList)

# Check for difference between stable and volatile groups
missingTest <- t.test(stableMissingData$numberMissed, volatileMissingData$numberMissed)

```

# Check that the success rate on the reaction time task waiting is approx. 50% #
``` {r RT, echo=FALSE}
#Basic RT success check for all participants
Hits = na.omit(FinalData$Press_hit)
AverageHits = mean(Hits)
sprintf('Overall success rate: %f', AverageHits)

#Check on success rates for each individual participant 
hitRates = c()
subj = c()
for(i in 1: length(CleanData)){
  thisData = CleanData[[i]]
  subj[i] = unique(thisData$participant)
  hit = na.omit(thisData$Press_hit)
  hitRates[i] = mean(hit)
}
hitRatesEval = data.frame(subj, hitRates)
hitRates = hitRates[!is.na(hitRates)]
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
Mode_hits = Mode(hitRates)

# Hit rates in stable condition 
StablehitRates = c()
Stablesubj = c()
for(i in 1: length(stableList)){
  thisData = stableList[[i]]
  Stablesubj[i] = unique(thisData$participant)
  hit = na.omit(thisData$Press_hit)
  StablehitRates[i] = mean(hit)
}
StablehitRates = StablehitRates[!is.na(StablehitRates)]
meanStablehit = mean(StablehitRates)

# Hit rates in volatile condition 
VolatilehitRates = c()
Volatilesubj = c()
for(i in 1: length(volatileGroup)){
  thisData = volatileList[[i]]
  Volatilesubj[i] = unique(thisData$participant)
  hit = na.omit(thisData$Press_hit)
  VolatilehitRates[i] = mean(hit)
}
VolatilehitRates = VolatilehitRates[!is.na(VolatilehitRates)]
meanVolatilehit = mean(VolatilehitRates)

#test the difference bewteen the two conditions 
hitRatesTest = t.test(StablehitRates, VolatilehitRates)

# Visualize the Rates
subj = seq(from=1,to=length(hitRates),by=1)
hitRatesdf = data.frame(subj, hitRates)
sortedHitRates = hitRatesEval[order(hitRatesEval$hitRates),]
# Remove instances when the rate is 0
zero = apply(hitRatesdf, 1, function(x) all(x !=0 ))
hitRatesdf = hitRatesdf[zero,]
#Create plot 
ggplot(data=hitRatesdf, aes(x=subj, hitRates)) + geom_point() + 
  theme_minimal() + labs(x="individuals", y="success rates")


```

# Start looking at basic checks of Data (mostly descriptive statistics) #
```{r descriptive, echo=FALSE}
#### All participants combined #####

#Average Precommit Total
pCommitTotal = mean(FinalData$Choice_Commit)
sprintf('Overall proportion Precommit: %f', pCommitTotal)

#Table for by the initial differences
print(data.frame(summarise(group_by(FinalData, All_difference=difference) , pCommit = mean(Choice_Commit))))
pCommitSub <- FinalData %>%
              group_by(participant, Condition) %>%
              summarise(pCommit = mean(Choice_Commit, na.rm = TRUE))

#### Participants Divided by Condition ####

#Average Precommit Total
pCommitVolatile = mean(volatileData$Choice_Commit)
medianVolatile = median(volatileData$Choice_Commit)
sprintf('Overall proportion Precommit in Volatile Condition: %f', pCommitVolatile)
pCommitStable = mean(stableData$Choice_Commit)
sprintf('Overall proportion Precommit in Stable Condition: %f', pCommitStable)

#Table for by the initial differences
#Volatile 
print(data.frame(summarise(group_by(volatileData, Volatile_difference=difference) , pCommit = mean(Choice_Commit))))
volDif = glm(Choice_Commit ~ difference, data=volatileData)
#Stable
print(data.frame(summarise(group_by(stableData, Stable_difference=difference) , pCommit = mean(Choice_Commit))))
staDif = glm(Choice_Commit ~ difference, data=stableData)
#Look at individual pCommits per group
#Volatile
pCommitSubVolatile = data.frame(summarise(group_by(volatileData, participant) , pCommit = mean(Choice_Commit)))
#Stable
pCommitSubStable = data.frame(summarise(group_by(stableData, participant) , pCommit = mean(Choice_Commit)))

### T test comparison of the two different conditions ###
#T-test
tTestCommit <- wilcox.test(volatileData$Choice_Commit, stableData$Choice_Commit, paired=FALSE)

#### Average Value Effect ### 
avGlm = glm(Choice_Commit ~ AverageValue, data=FinalData)
avStable = glm(Choice_Commit ~ AverageValue, data=stableData)
avVolatile = glm(Choice_Commit ~ AverageValue, data=volatileData)

```

# Look at learning the task structure over time #
```{r learning, echo=FALSE}
# Function to extract specific rows from a data frame
extract_rows <- function(df, rows) {
  df[rows, ]
}

# Extracting the first and last blocks for each group
first_rows <- 1:25
last_rows <- 73:96

firstStable <- lapply(stableList, extract_rows, rows = first_rows)
firstVolatile <- lapply(volatileList, extract_rows, rows = first_rows)
lastStable <- lapply(stableList, extract_rows, rows = last_rows)
lastVolatile <- lapply(volatileList, extract_rows, rows = last_rows)

#Combine the lists into different dataframes 
#volatile - first 
volatilefirstCombined <- do.call("rbind", firstVolatile)
volatileFirstData <- volatilefirstCombined[volatilefirstCombined$Choice_Miss == F,]
#volatile - last 
volatilelastCombined <- do.call("rbind", lastVolatile)
volatileLastData = volatilelastCombined[volatilelastCombined$Choice_Miss == F,]
#stable - first 
stablefirstCombined <- do.call("rbind", firstStable)
stableFirstData <- stablefirstCombined[stablefirstCombined$Choice_Miss == F,]
#stable - last 
stablelastCombined <- do.call("rbind", lastStable)
stableLastData <- stablelastCombined[stablelastCombined$Choice_Miss == F,]

# Calculating pCommit for each group
calc_pCommit <- function(data) mean(data$Choice_Commit)
pCommit_FirstVolatile <- calc_pCommit(volatileFirstData)
pCommit_FirstStable <- calc_pCommit(stableFirstData)
pCommit_LastVolatile <- calc_pCommit(volatileLastData)
pCommit_LastStable <- calc_pCommit(stableLastData)

# Calculate differences
diffVolatile <- pCommit_LastVolatile - pCommit_FirstVolatile
diffStable <- pCommit_LastStable - pCommit_FirstStable

# Summarize groups 
summarize_group <- function(data) {
  dplyr::summarise(dplyr::group_by(data, participant), pCommit = mean(Choice_Commit))
}

volatileFirstpCommit <- summarize_group(volatileFirstData)
stableFirstpCommit <- summarize_group(stableFirstData)
volatileLastpCommit <- summarize_group(volatileLastData)
stableLastpCommit <- summarize_group(stableLastData)

#See if significant differences 
wilcoxVolatile = wilcox.test(volatileFirstpCommit$pCommit, volatileLastpCommit$pCommit, paired = TRUE)
wilcoxStable = wilcox.test(stableFirstpCommit$pCommit, stableLastpCommit$pCommit, paired = TRUE)

#Print the findings
sprintf('pCommit first Volatile: %f', pCommit_FirstVolatile)
sprintf('pCommit first Stable: %f', pCommit_FirstStable)
sprintf('pCommit last Volatile: %f', pCommit_LastVolatile)
sprintf('pCommit last Stable: %f', pCommit_LastStable)
sprintf('Last - First Volatile: %f', diffVolatile)
sprintf('Last - First Stable: %f', diffStable)

# Preparing data for plots
prepare_plot_data <- function(firstData, lastData, firstBlockNum, lastBlockNum) {
  firstPar <- summarize_group(firstData)
  firstPar$block <- rep(firstBlockNum, nrow(firstPar))
  lastPar <- summarize_group(lastData)
  lastPar$block <- rep(lastBlockNum, nrow(lastPar))
  rbind(firstPar, lastPar)
}

stable_df <- prepare_plot_data(stableFirstData, stableLastData, 1, 4)
volatile_df <- prepare_plot_data(volatileFirstData, volatileLastData, 1, 4)

# Plotting code
# plot stable change from first to last block 
stable_change_plot = ggplot()+
  geom_point(data=stable_df, aes(x=block, y=pCommit, group=participant), color="lightblue")+ 
  geom_line(data=stable_df, aes(x=block, y=pCommit, group=participant), color="lightblue")+ 
  theme_minimal() + labs(x="Block", y="Proportion Precommit", title="Stable")
# plot volatile change from first to last block 
volatile_change_plot = ggplot()+
  geom_point(data=volatile_df, aes(x=block, y=pCommit, group=participant), color="darkblue")+ 
  geom_line(data=volatile_df, aes(x=block, y=pCommit, group=participant), color="darkblue")+ 
  theme_minimal() + labs(x="Block", y="Proportion Precommit", title="Volatile")
#Plot of both Stable and Volatile change
full_change_plot = ggplot()+
  geom_point(data=volatile_df, aes(x=block, y=pCommit, group=participant), color="darkblue")+ 
  geom_line(data=volatile_df, aes(x=block, y=pCommit, group=participant), color="darkblue")+ 
  geom_point(data=stable_df, aes(x=block, y=pCommit, group=participant), color="lightblue")+ 
  geom_line(data=stable_df, aes(x=block, y=pCommit, group=participant), color="lightblue")+ 
  theme_minimal() + labs(x="Block", y="Proportion Precommit", title="")
```

# Explore model of best fit with data #
```{r models, echo=FALSE}

#### All data combined #### 

# First test if condition had an effect 
groupModel = glmer(Choice_Commit ~  Condition  + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
# full model 
fullModel = glmer(Choice_Commit ~ centeredAverageValue * trial * centeredDifference  + (1 | participant), data= FinalData, family=binomial, nAGQ = 0)

# Trial only
fullTrial = glmer(Choice_Commit ~  trial  + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
# Trial and condition 
fullTrialCondition = glmer(Choice_Commit ~  trial  * Condition + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)

# Average Value only 
fullAverageValue = glmer(Choice_Commit ~  centeredAverageValue  + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)

# Difference Only 
fullDifference = glmer(Choice_Commit ~  centeredDifference  + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)

# Interaction between difference and average value
FullFirstinteraction = glmer(Choice_Commit ~ centeredAverageValue + centeredDifference + centeredAverageValue * centeredDifference  + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
# Comparison 
anovaFull = anova(fullDifference, fullTrial, fullAverageValue, FullFirstinteraction, fullModel)

#### Separate Stable and Volatile Conditions #### 

# Full 
stableFull = glmer(Choice_Commit ~ centeredAverageValue * trial * centeredDifference  + (1 | participant), data= stableData, family=binomial, nAGQ = 0)
volatileFull = glmer(Choice_Commit ~ centeredAverageValue * trial * centeredDifference  + (1 | participant), data= volatileData, family=binomial, nAGQ = 0)

# Trial only
stableTrial = glmer(Choice_Commit ~  trial  + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
volatileTrial = glmer(Choice_Commit ~  trial  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)

# Average Value only 
stableAverageValue = glmer(Choice_Commit ~  centeredAverageValue  + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
volatileAverageValue = glmer(Choice_Commit ~  centeredAverageValue  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)


# Difference only 
stableDifference = glmer(Choice_Commit ~  centeredDifference  + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
volatileDifference = glmer(Choice_Commit ~  centeredDifference  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)

# Check on the interactions 
stableFirstinteraction = glmer(Choice_Commit ~ centeredAverageValue + centeredDifference + centeredAverageValue * centeredDifference  + (1  | participant),
                               data= stableData, family=binomial, nAGQ = 0)
volatileFirstinteraction = glmer(Choice_Commit ~ centeredAverageValue + centeredDifference + centeredAverageValue * centeredDifference  + (1  | participant), 
                               data= volatileData, family=binomial, nAGQ = 0)

#Comparison 
anovaStable = anova(stableDifference, stableTrial, stableAverageValue, stableFirstinteraction, stableFull)
anovaVolatile = anova(volatileDifference, volatileTrial, volatileAverageValue, volatileFirstinteraction, volatileFull)

```

# Complete full model comparison #
```{r model_comp, echo=FALSE}

## Start with full model ##
full_1  = glmer(Choice_Commit ~  Condition  + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
full_2 = glmer(Choice_Commit ~  Condition  + trial + Condition * trial + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
full_3 = glmer(Choice_Commit ~  Condition  + trial + centeredDifference + Condition * trial + Condition * centeredDifference + centeredDifference * trial + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
full_4 = glmer(Choice_Commit ~  Condition  + trial + centeredAverageValue + Condition * trial + Condition * centeredAverageValue + centeredAverageValue * trial + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
full_5 = glmer(Choice_Commit ~  Condition  + trial + centeredAverageValue + centeredDifference + Condition * trial + Condition * centeredAverageValue + Condition * centeredDifference + centeredAverageValue * trial + centeredDifference * trial + centeredDifference * centeredAverageValue + (1  | participant), data= FinalData, family=binomial, nAGQ = 0)
anovaFull = anova(full_1, full_2, full_3, full_4, full_5)

## Model comparison for Stable ##
stable_1 = glmer(Choice_Commit ~  trial  + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
stable_2 = glmer(Choice_Commit ~  centeredDifference  + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
stable_3 = glmer(Choice_Commit ~  centeredAverageValue  + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
stable_4 = glmer(Choice_Commit ~  trial  + centeredDifference + trial * centeredDifference + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
stable_5 = glmer(Choice_Commit ~  trial  + centeredAverageValue + trial * centeredAverageValue + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
stable_6 = glmer(Choice_Commit ~  trial  + centeredDifference + centeredAverageValue + trial * centeredDifference  + trial * centeredAverageValue + centeredDifference * centeredAverageValue + (1  | participant), data= stableData, family=binomial, nAGQ = 0)
anovaStable = anova(stable_1, stable_2, stable_3, stable_4, stable_5, stable_6)
  
## Model comparison for Volatile ##
volatile_1 = glmer(Choice_Commit ~  trial  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)
volatile_2 = glmer(Choice_Commit ~  centeredDifference  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)
volatile_3 = glmer(Choice_Commit ~  centeredAverageValue  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)
volatile_4 = glmer(Choice_Commit ~  trial  + centeredDifference + trial * centeredDifference + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)
volatile_5 = glmer(Choice_Commit ~  trial  + centeredAverageValue + trial * centeredAverageValue + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)
volatile_6 = glmer(Choice_Commit ~  trial  + centeredDifference + centeredAverageValue + trial * centeredDifference  + trial * centeredAverageValue + centeredDifference * centeredAverageValue + (1  | participant), data= volatileData, family=binomial, nAGQ = 0)
anovaVolatile = anova(volatile_1, volatile_2, volatile_3, volatile_4, volatile_5, volatile_6)

```

# Fit models to individual participants #
```{r ind_models, echo=FALSE}
# Model for Volatile -> Choice_Commit ~ centeredAverageValue * trial * centeredDifference  + (1  | participant), data= volatileData, family=binomial, nAGQ = 0
# model for Stable ->  Choice_Commit ~ centeredAverageValue * trial * centeredDifference  + (1  | participant), data= stableData, family=binomial, nAGQ = 0

allBetas = list()
for (i in 1: length(CleanData)){
  d = CleanData[[i]]
  model = lm(Choice_Commit ~ centeredAverageValue + trial + centeredDifference + centeredAverageValue * trial + centeredAverageValue * centeredDifference + centeredDifference * trial, data=d)
  coef = data.frame(model$coefficients)
  coef = data.frame(t(coef))
  par = unique(d$participant)
  coef = data.frame(coef, par)
  allBetas[[i]] = coef
}
allBetasCombined = do.call("rbind", allBetas)

# Edit the Volatile group for the participant without the qualtrics data 
volatileGroupCor = volatileList[-33]
# Individual Coefficients for Volatile
volatileBetas = list()
for (i in 1: length(volatileGroupCor)){
  d = volatileGroupCor[[i]]
  model = lm(Choice_Commit ~ centeredAverageValue * trial * centeredDifference, data=d)
  coef = data.frame(model$coefficients)
  coef = data.frame(t(coef))
  volatileBetas[[i]] = coef
}
allVolatileBetas = do.call("rbind", volatileBetas)

# Individual Coefficients for Stable
stableBetas = list()
for (i in 1: length(stableList)){
  d = stableList[[i]]
  model = lm(Choice_Commit ~ centeredAverageValue * trial * centeredDifference, data=d)
  coef = data.frame(model$coefficients)
  coef = data.frame(t(coef))
  stableBetas[[i]] = coef
}
allStableBetas = do.call("rbind", stableBetas)


## Simple T test comparisons ##
#Not significant 
t.test(allStableBetas$X.Intercept., allVolatileBetas$X.Intercept.)
t.test(allStableBetas$centeredAverageValue, allVolatileBetas$centeredAverageValue)
t.test(allStableBetas$centeredDifference, allVolatileBetas$centeredDifference)
t.test(allStableBetas$centeredAverageValue.trial, allVolatileBetas$centeredAverageValue.trial)
t.test(allStableBetas$centeredAverageValue.centeredDifference, allVolatileBetas$centeredAverageValue.centeredDifference)
t.test(allStableBetas$centeredAverageValue.trial.centeredDifference, allVolatileBetas$centeredAverageValue.trial.centeredDifference)

#Significant 
t.test(allStableBetas$trial, allVolatileBetas$trial)
t.test(allStableBetas$trial.centeredDifference, allVolatileBetas$trial.centeredDifference)

```

# Load questionnaire data #
```{r load_questionnaires, echo=FALSE}
## Set directory to get questionnaire data ##
questionnaire = read.csv('~/Documents/Github/precommitment/Exp_2_questionnaire.csv')

## Number with missing data ##
numMissing = as.numeric(complete.cases(questionnaire))
# Change that one participant 
questionnaire[98,1] = 1100
questionnaire[159,1] = 1163
questionnaire[106,1] = 1108
#Remove participant with missing data 
questionnaire = questionnaire[-c(54, 68, 71, 109, 167),]

## Participant ID variable ##
id = questionnaire$Q1
pCommitQual = pCommitSub[-71,]
```

# BIS-11 variables #
```{r BIS, echo=FALSE}
# Function to replace missing values with median
replace_missing_with_median <- function(data) {
  for (i in 1:nrow(data)) {
    data[i,][is.na(data[i,])] <- median(na.omit(data[i,]))
  }
  data
}

# BIS Data Preparation
bisdata <- questionnaire[, 31:60]
keys <- c(-1,1,1,1,1,1,-1,-1,-1,-1,1,-1,-1,1,-1,1,1,1,1,-1,1,1,1,1,1,1,1,1,-1,-1)
clean_bis <- reverse.code(keys, bisdata, mini=1, maxi=4)
clean_bis <- replace_missing_with_median(clean_bis)

# Function to calculate sum and sort
calculate_and_sort <- function(data, id, columns) {
  sum_data <- rowSums(data[, columns])
  sorted_data <- data.frame(id, sum_data)[order(id),]
}

# Calculating sums
participantBisSum <- calculate_and_sort(clean_bis, id, 1:ncol(clean_bis))
participantAttentional <- calculate_and_sort(clean_bis, id, c(5,9,11,20,28,6,24,26))
participantMotor <- calculate_and_sort(clean_bis, id, c(2,3,4,17,19,22,25,16,21,23,30))
participantNonplanning <- calculate_and_sort(clean_bis, id, c(1,7,8,12,13,14,10,15,18,27,29))

# Creating Full_BIS DataFrame
Full_BIS <- data.frame(Bis_score = participantBisSum, Attentional = participantAttentional$sum_data, 
                       Motor = participantMotor$sum_data, Nonplanning = participantNonplanning$sum_data)
dfBIS = data.frame(pCommitQual, Full_BIS)
# Separate into the two different groups 
bisStable = dfBIS[which(dfBIS$Condition=="Stable"),]
bisVolatile = dfBIS[which(dfBIS$Condition=="Volatile"),]

# Function for Correlation Computation
calculate_correlations <- function(data, variable) {
  expression_text <- paste("cor.test(data$pCommit, data$", variable, ")", sep = "")
  eval(parse(text = expression_text))
}

# Computing Correlations
Full_correlations <- lapply(names(Full_BIS)[-1], function(var) calculate_correlations(dfBIS, var))
Stable_correlations <- lapply(names(Full_BIS)[-1], function(var) calculate_correlations(bisStable, var))
Volatile_correlations <- lapply(names(Full_BIS)[-1], function(var) calculate_correlations(bisVolatile, var))

```


